The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2023-05-01 MS_ToSubmit/MS_Burdwood_foodweb.tex
Updates Claudia Boy

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017). In addition,
such MPAs are part of a network of protected areas in the sub-Antarctic
area (jointly with MPA Yaganes) that aims to protect and strengthen the
functioning of this southern region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
L. Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; L. Schejter \& Bremec, 2019; Sirito, 2019), and has an important
role in the life history of fishes as a food source, refuge and nursery
area (Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al.,
2022; Florencia et al., 2023; García Alonso et al., 2018; Troccoli et
al., 2020). The maintenance of this singular community is related to
local and regional oceanographic processes, including the circulation of
the rich Malvinas (Falkland) current in the area (Guerrero et al., 1999;
A. R. Piola \& Gordon, 1989) and the upwelling and mixing phenomena
(Matano et al., 2019). The input of nutrients from the Malvinas
(Falkland) current also supports a diverse plankton community (Guinder
et al., 2020). Overall, 811 benthic and plankton species have been
identified for the MPA N-BB ecosystem, where 349 were reported for the
first time in the area in recent years (APN, 2022). However, identifying
the main species involved in the maintenance of ecosystem services and
health as well as for management and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure than other sub-Antarctic areas, such as the Atlantic coast of
Tierra del Fuego. The Fuegian sprat \emph{Sprattus fuegensis} and
longtail southern cod \emph{Patagonotothen ramsayi} are considered the
most plausible `wasp-waist' species (Riccialdelli et al., 2020).

High-latitude marine ecosystems, such as the MPA N-BB, are characterized
by an unexpected complexity in terms of biodiversity and ecological
interactions (Cordone et al., 2020; Day et al., 2013; Kortsch et al.,
2019; Trathan et al., 2021). Although there is a robust knowledge about
the complexity considering the richness of the benthic and plankton
communities in the MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020;
L. Schejter et al., 2016, 2020), a better understanding of species
interactions' complexity and structure is needed. This aspect can be
tackled by analyzing one of the most frequent relationships between
species: the predator-prey interaction (Bascompte, 2009). The sum of
predator-prey or trophic interactions of a particular region is referred
to as a food web, representing the roadmap for matter and energy flow in
an ecosystem. In recent years, network approaches have been successfully
applied to study complex high-latitude marine ecosystems, improving our
knowledge on structure, functioning, and response to
environmental/anthropogenic changes (Cordone et al., 2018; Funes et al.,
2022; Kortsch et al., 2015; Marina et al., 2023). Among anthropogenic
threats, it is worth mentioning that contaminants like mercury and
microplastics have been recently reported as important threats to the
MPA N-BB region (Cossi et al., 2021; Di Mauro et al., 2022; Fioramonti
et al., 2022).

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023). They are located 150 km east of Isla de
los Estados and 200 km south of Malvinas/Falkland Islands (Figure 1).
The MPA N-BB I comprises nearly 34,000 km\textsuperscript{2}
circumscribed by the 200 m isobath, between 54º--55ºS and 56º--62ºW,
with a slight slope extended nearly 370 km east--west. Physical features
in the BB are fairly stable, with salinity averaging 34 all year round
and temperature ranging between 4 and 8ºC (Acha et al., 2004; Guerrero
et al., 1999; A. Piola \& Falabella, 2009). The BB is surrounded by
steep flanks of up to 4000 m depth through which strong currents
circulate (Matano et al., 2019; A. R. Piola \& Gordon, 1989; Reta,
2014). Such deep slope forms the basis of N-BB II, protecting about
32,000 km\textsuperscript{2}. Intense upwelling and mixing occur over
it, entraining deep nutrient-rich waters into the photic layer (Matano
et al., 2019; A. Piola \& Falabella, 2009) and resulting in a fairly
homogeneous water column both spatially and temporally (Glorioso \&
Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

@@

In order to build the network of predator-prey interactions, we reviewed
more than 150 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to xx research cruises conducted
in the MPAs N-BB I and II. Here it is important to note that the
sampling effort was greater in the MPA N-BB I. Furthermore, we took into
account personal communications from experts belonging to the working
group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

@@

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated most
of them (Figure 3A). The distribution of prey among predators showed
that most consumers fed on a low number of prey whereas few had multiple
preys (Figure 3B). These were the top-five predators in number of prey:
yellowfin notothen \emph{Patagonotothen guntheri} (Notothenioid fish, 52
prey), rock cod \emph{Patagonotothen ramsayi} (Notothenioid fish, 50
prey), Patagonian toothfish \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), broad nose skate \emph{Bathyraja
brachyurops} (Chondrichthyan, 30 prey), and graytail skate
\emph{Bathyraja griseocauda} (Chondrichthyan, 28 prey). Following the
same distribution pattern, few prey presented multiple predators (Figure
3C). These were the top-five prey in number of predators: Detritus
(Non-living, 153 predators), the three categories of Diatoms considered
(benthic, centric and pennate, 72 predators on average), and species of
the genus \emph{Euphausia} (Zooplankton, 46 predators). Finally, taking
into account the interactions within each functional group, again most
interactions were concentrated in a few species (Figure 3D). The most
evident species were: \emph{Themisto gaudichaudii} (Amphipoda),
\emph{Zygochlamys patagonica} (Bivalvia), \emph{Aspidostoma giganteum}
(Bryozoa), \emph{Munida gregaria} (Decapoda), \emph{Patagonotothen
ramsayi} and \emph{Patagonotothen guntheri} (bentho-pelagic fish),
\emph{Psychrolutes marmoratus} (demersal fish), and species of
\emph{Euphausia} (Zooplankton). Overall, there is an evident asymmetry
in the distribution of interactions among species at different levels in
the MPA N-BB food web.

@@

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). The
most evident significant relationship was with trophic similarity,
i.e.~the higher the species' TL, the lower the trophic similarity or the
higher the uniqueness in terms of trophic role (Figure 4C). Here it is
noteworthy to highlight those high-trophic level species with low values
of trophic similarity: \emph{Bathyraja macloviana} and \emph{Squalus
acanthias} (Chondrichthyans), \emph{Diplopteraster clarki} and
\emph{Pteraster sp} (echinoderms), \emph{Phalacrocorax\_atriceps} and
\emph{Eudyptes chrysocome} (sea birds), and \emph{Lagenorhynchus
cruciger} and \emph{Mesoplodon bowdoini} (marine mammals) (Table S3).

@@

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.77); `module hub' was
constrained to mid-TL species (2.48 - 3.92); `module connector' from low
to mid-TLs (1 - 3.86); and `network connector', the least frequent, had
all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D,
see Figure S2 for species' topological roles in a food web graph
framework). Here it is important to highlight the two latter topological
roles because they are responsible for linking modules and maintaining
the connectivity of the food web: 40 species (5 network connectors + 35
module connectors) from 20 different functional groups with a TL range =
1 - 3.86. The following were the five species with a network connector
role: Detritus, Diatoms (benthic, centric and pennate), and
\emph{Patagonotothen ramsayi}. The 35 species with a module connector
role represent these functional groups: Amphipoda, Bivalvia,
Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, detritus, Diatoms,
Dinoflagellates, Echinodermata, fish (bentho-pelagic, demersal,
Chondrichthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, sea
birds, and zooplankton (see Supplementary Material Table S3 for the
identity of the species).

@@

\caption{Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 72.48x - 111.98, R^2 = 0.04, p-value < 3.38e-05$), closeness ($y = 5.78e-06x - 9.37e-04, R^2 = -0.0005, p-value = 0.37$) and trophic similarity ($y = -0.01x - 0.11, R^2 = 0.07, p-value = 6.76e-08$). Note that for A, B and C panels only species with TLs equal or greater than 2 were considered.}\label{fig:figure4}
\end{figure}

@@

The degree distribution, the distribution of the number of interactions
per species, is the core of the structure of species interactions, which
influences the opportunities for multiple species to persist in the long
term and, therefore, their coexistence (Godoy et al., 2018). The food
web for the MPA N-BB presents an asymmetric degree distribution. This
pattern was identified at different levels of analysis: food web,
predator, prey, and functional group. Such asymmetry is a well-known
feature in empirical complex food webs in particular (Dunne et al.,
2002a; Montoya \& Solé, 2003; Stouffer et al., 2005), and has received
great attention in complex networks in general (Albert \& Barabási,
2002; Newman, 2003). The degree distribution affects the resilience of
complex food webs against random failures and intentional attacks: food
webs showing right-skewed distributions, like the one described in this
study, are more vulnerable to the removal of the most connected species
or hubs, with the potential of producing secondary extinctions and a
catastrophic fragmentation of the network (Albert et al., 2000; Dunne et
al., 2002b; Eklöf \& Ebenman, 2006).

@@

On the other hand, a low number of prey are consumed by many predators
in the food web of the MPA N-BB. This suggests that there are dominant
food sources from which most consumers depend on and from where the
ecosystem energy is being transferred to the upper trophic levels. The
most demanded source we identified in this study (i.e.~detritus)
supports the abundant benthic community of filter-feeders (L. Schejter
et al., 2016). Furthermore, we found that the second and third-most
consumed prey were diatoms and species of \emph{Euphausia},
respectively, which are essential sources for the diverse zooplankton
community (Spinelli et al., 2020), mid-TL consumers like the Fuegian
sprat \emph{Sprattus fuegensis} (Padovani et al., 2021) and
\emph{Patagonotothen ramsayi} (Fischer et al., 2022), and top predators
(Raya Rey, 2022; Riccialdelli, 2022).

@@

Closeness and betweenness are defined as ``mesoscale'' properties
because they consider direct and indirect interactions, therefore
describing the focal species' ability to influence the rest of the
species of the food web (Lai et al., 2012). Closeness quantifies how
many steps away species \(i\) is from all other species in the food web,
and is proportional to how rapidly the indirect effects of the focal
species can spread to other species in the network (Scotti \& Jordán,
2010). In the food web of the MPA N-BB, low-TL consumers arise as
important in this regard: species of the zooplankton community,
\emph{Calanus} and \emph{Euphausia}, and Brachiopoda. Any perturbation
affecting these species, such as the recently confirmed contaminants
mercury (Fioramonti et al., 2022) and microplastics (Cossi et al., 2021;
Di Mauro et al., 2022) for the area, should be of concern since it might
reach many other species in the food web. Otherwise, betweenness
measures the number of shortest paths between species, providing
information on the importance of species as ``bridges'' for energy
transfer: a species with high betweenness takes part in more food chains
and therefore affects more energy flows (Scotti \& Jordán, 2010). We
have identified the longtail southern cod \emph{Patagonotothen ramsayi}
as the most important species in this sense. Moreover, in light of our
analysis, other species like the Patagonian toothfish \emph{Dissostichus
eleginoides}, the Patagonian cod \emph{Salilota australis}, the
yellowfin notothen \emph{Patagonotothen guntheri}, and the Patagonian
longfin squid \emph{Doryteuthis gahi} should be considered as relevant
in the energy transfer in the ecosystem. All these species have a
mid-trophic position in the food web, supporting the `wasp-waist'
control hypothesis for the MPA N-BB (Riccialdelli et al., 2020).

Ecosystems with a pronounced `wasp-waist' structure are suggested to
present a high trophic redundancy, since many species would show similar
trophic habits (Cury et al., 2000). The significant negative
relationship between trophic similarity and trophic level enhances the
hypothesis of functional redundancy at low and mid-TL species compared
to higher TL species for the MPA N-BB food web (Riccialdelli et al.,
2020). At the same time, our results highlight the uniqueness in terms
of the trophic role of high-TL predators. Here, not only the expected
pelagic animals such as marine mammals and sea birds arise as relevant,
but also benthic vertebrate (chondrichthyans \emph{Bathyraja macloviana}
and \emph{Squalus acanthias}) and invertebrate species (echinoderms
\emph{Diplopteraster clarki} and \emph{Pteraster sp}) are noteworthy.
The role that such species play in the MPA N-BB ecosystem is unique and
perturbations on them might result in unprecedented changes at the
trophic structure and functioning level.

@@

Caveats: The food web studied in the present work might be more
representative of the MPA Namuncurá - Banco Burdwood I ecosystem, which
comprises the shallow submarine plateau called Burdwood Bank. This is
related to the sampling effort that was conducted during the research
cruises in the former MPA compared to the MPA N-BB II (i.e.~deep flanks
to the south). As a consequence most of the data we used to build the
network comes from studies performed in the MPA N-BB I. Despite this
fact, we decided to build the food web considering both MPAs due to the
tight oceanographic and ecological connection that exists among them
(references in APN, 2022).

It's worth mentioning that we did not consider quantitative data
(i.e.~abundance, biomass) to assess the species' role in the food web.
Although there exists such data for some species of the MPA N-BB
(Schejter \& Albano, 2021), it would not be possible to incorporate it
in the food web framework described here due to a taxonomical resolution
mismatch. In this regard, we should mention the case of
\emph{Zygochlamys patagonica} (Bivalvia) that is highlighted by our
species-level analyses though it has been found in low abundances in the
area (Schejter \& Albano, 2021).

Some species of sessile suspension feeders in high-latitude marine
ecosystems, such as sponges, ascidians and octocorals, avoid predation
by producing secondary metabolites that function as a chemical defense
(Moles et al., 2015; Núñez-Pons et al., 2010; Prieto et al., 2022).
Although this was not reported at the MPA N-BB, there are few studies
that reported it other locations in species that were found in the MPA
N-BB (Rojo de Almeida et al., 2010). Thus, the interactions that we
considered between these species and their predators should be taken
with caution.

Seasonal variation in predators' diet due to food source availability.
Brose et al. (2019)

Spatial variation of study area: Banco Burdwood west-east gradient. In
the oceanic BB area, complex oceanographic conditions generate an
internal spatial heterogeneity, mainly along its longitudinal axis
(Matano et al., 2019), and consequently, plankton communities respond
with different spatial distributions (Spinelli et al., 2020). Kortsch et
al. (2019)

@@

\begin{itemize}
\item
  Incorporate habitat heterogeneity, since the benthic community in the
  BB and deep flanks differs (Schejter \& Albano, 2021). This
  heterogeneity might lead to distinct food web properties, hence
  differences in ecosystem functioning and response to anthropogenic and
  environmental perturbations (Cordone et al., 2020). El caso de las
  ascidias (Tatián)
\item
  Incorporate species traits, like body size and mass, which are known
  to be important drivers in predator-prey interactions.
\item
  Simulate perturbations on food web and target species: mercury
  transfer, microplastic pollution, fisheries. Cossi et al. (2021); Di
  Mauro et al. (2022); Fioramonti et al. (2022)
\item
  Estimation of interaction strength considering species and interaction
  traits (i.e.~body size, body mass, interaction dimensionality) to gain
  insights into food web stability and response to anthropogenic and
  environmental perturbations. Nilsson \& McCann (2016); Cirtwill et al.
  (2018); Kortsch et al. (2021)
\end{itemize}

@@

>>>>

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá - Banco
Burdwood I and II (MPA N-BB, 53º--55ºS 56º--62ºW), which is the focus of
this work, is unique since these MPAs are associated to a submarine
plateau and its southern adjacent deep slope region, respectively
(Falabella, 2017). In addition, such MPAs are part of a network of
protected areas in the sub-Antarctic area (jointly with MPA Yaganes)
that aims to protect and strengthen the functioning of this southern
region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources, 2009). Also, it
provides habitat to several small-sized species (López-Gappa et al.,
2018; Schejter \& Bremec, 2019; Sirito, 2019), and has an important role
in the life history of fishes as a food source, refuge and nursery area
(Covatti Ale et al., 2022; Delpiani et al., 2020; Fischer et al., 2022;
Florencia et al., 2023; García Alonso et al., 2018; Troccoli et al.,
2020). The maintenance of this singular community is related to local
and regional oceanographic processes, including the circulation of the
rich Malvinas (Falkland) current in the area (Guerrero et al., 1999; A.
R. Piola \& Gordon, 1989) and the upwelling and mixing phenomena (Matano
et al., 2019). The input of nutrients from the Malvinas (Falkland)
current also supports a diverse plankton community (Guinder et al.,
2020). Overall, 811 benthic and plankton species have been identified
for the MPA N-BB ecosystem, where 349 were reported for the first time
in the area in recent years (APN, 2022). However, identifying the main
species involved in the maintenance of ecosystem services and health as
well as for management and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure, meaning a shorter food-chain length and a greater trophic
overlap and redundancy, than other sub-Antarctic areas, such as the
Atlantic coast of Tierra del Fuego. The Fuegian sprat \emph{Sprattus
fuegensis} and longtail southern cod \emph{Patagonotothen ramsayi} are
considered the most plausible `wasp-waist' species (Riccialdelli et al.,
2020).

High-latitude marine ecosystems, such as the MPA N-BB, are considered
complex systems in terms of biodiversity and ecological interactions
(Cordone et al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan
et al., 2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (APN, 2022; Guinder et al., 2020; Schejter et al.,
2016, 2020), a better understanding of species interactions' complexity
and structure is needed. This aspect can be tackled by analyzing one of
the most frequent relationships between species: the predator-prey
interaction (Bascompte, 2009). The sum of predator-prey or trophic
interactions of a particular region is referred to as a food web,
representing the roadmap for matter and energy flow in an ecosystem. In
recent years, network approaches have been successfully applied to study
complex high-latitude marine ecosystems, improving our knowledge on
structure, functioning, and response to environmental/anthropogenic
changes (Cordone et al., 2018; Funes et al., 2022; Kortsch et al., 2015;
Marina et al., 2023). Among anthropogenic threats, it is worth
mentioning that contaminants like mercury and microplastics have been
recently reported as important threats to the MPA N-BB region (Cossi et
al., 2021; Di Mauro et al., 2022; Fioramonti et al., 2022).

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws
26.875 and 27.490, comprise a shallow submarine plateau called Burdwood
Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB
II, respectively (APN, 2023) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 34,000 km\textsuperscript{2} circumscribed
by the 200 m isobath, between 54º--55ºS and 56º--62ºW, with a slight
slope extended nearly 370 km east--west. Physical features in the BB are
fairly stable, with salinity averaging 34 all year round and temperature
ranging between 4 and 8ºC (Acha et al., 2004; Guerrero et al., 1999; A.
Piola \& Falabella, 2009). The BB is surrounded by steep flanks of up to
4000 m depth through which strong currents circulate (Matano et al.,
2019; A. R. Piola \& Gordon, 1989; Reta, 2014). Such deep slope forms
the basis of N-BB II, protecting about 32,000 km\textsuperscript{2}.
Intense upwelling and mixing occur over it, entraining deep
nutrient-rich waters into the photic layer (Matano et al., 2019; A.
Piola \& Falabella, 2009) and resulting in a fairly homogeneous water
column both spatially and temporally (Glorioso \& Flather, 1995;
Guerrero et al., 1999; Matano et al., 2019).

@@

In order to build the network of predator-prey interactions, we reviewed
more than 150 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to 16 research cruises conducted
in the MPAs N-BB I and II during 2014-2019. Here it is important to note
that the sampling effort was greater in the MPA N-BB I. Furthermore, we
took into account personal communications from experts belonging to the
working group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

@@

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated most
of them (Figure 3A). The distribution of prey among predators showed
that most consumers fed on a low number of prey whereas few had multiple
preys (Figure 3B). The top-five predators in number of prey were:
yellowfin notothen \emph{Patagonotothen guntheri} (Notothenioid fish, 52
prey), rock cod \emph{Patagonotothen ramsayi} (Notothenioid fish, 50
prey), Patagonian toothfish \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), broad nose skate \emph{Bathyraja
brachyurops} (Chondrichthyan, 30 prey), and graytail skate
\emph{Bathyraja griseocauda} (Chondrichthyan, 28 prey). Following the
same distribution pattern, few prey presented multiple predators (Figure
3C). The top-five prey in number of predators were: Detritus
(Non-living, 153 predators), the three categories of Diatoms considered
(benthic, centric and pennate, 72 predators on average), and species of
the genus \emph{Euphausia} (Zooplankton, 46 predators). Finally, taking
into account the interactions within each functional group, most
interactions were concentrated in a few species (Figure 3D). The most
evident species were: \emph{Themisto gaudichaudii} (Amphipoda),
\emph{Zygochlamys patagonica} (Bivalvia), \emph{Aspidostoma giganteum}
(Bryozoa), \emph{Munida gregaria} (Decapoda), \emph{Patagonotothen
ramsayi} and \emph{Patagonotothen guntheri} (bentho-pelagic fish),
\emph{Psychrolutes marmoratus} (demersal fish), and species of
\emph{Euphausia} (Zooplankton). Overall, there is an evident asymmetry
in the distribution of interactions among species at different levels in
the MPA N-BB food web.

@@

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). The
most evident significant relationship was with trophic similarity,
i.e.~the higher the species' TL, the lower the trophic similarity or the
higher the uniqueness in terms of trophic role (Figure 4C). Here it is
noteworthy to highlight those high-trophic level species with low values
of trophic similarity: \emph{Bathyraja macloviana} and \emph{Squalus
acanthias} (Chondrichthyans), \emph{Diplopteraster clarki} and
\emph{Pteraster sp} (echinoderms), \emph{Phalacrocorax atriceps} and
\emph{Eudyptes chrysocome} (sea birds), and \emph{Lagenorhynchus
cruciger} and \emph{Mesoplodon bowdoini} (marine mammals) (Table S3).

@@

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.77); `module hub' was
constrained to mid-TL species (2.48 - 3.92); `module connector' from low
to mid-TLs (1 - 3.86); and `network connector', the least frequent, had
all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D,
see Figure S2 for species' topological roles in a food web graph
framework). Here it is important to highlight the two latter topological
roles because they are responsible for linking modules and maintaining
the connectivity of the food web: 40 species (5 network connectors + 35
module connectors) from 20 different functional groups with a TL range =
1 - 3.86. The following were the five species with a network connector
role: Detritus, Diatoms (benthic, centric and pennate), and
\emph{Patagonotothen ramsayi}. The 35 species with a module connector
role represent these functional groups: Amphipoda, Bivalvia,
Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, detritus, Diatoms,
Dinoflagellates, Echinodermata, fish (bentho-pelagic and demersal
Osteichthyes, and Chondrychthyes), Foraminifera, Polychaeta, Porifera,
Pycnogonida, sea birds, and zooplankton (see Supplementary Material
Table S3 for the identity of the species).

@@

\caption{Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = x - , R^2 = , p-value < $), closeness ($y = x - , R^2 = , p-value = $) and trophic similarity ($y = x - , R^2 = , p-value = $). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered.}\label{fig:figure4}
\end{figure}

@@

The degree distribution, the distribution of the number of interactions
per species, is the core of the structure of species interactions, which
influences the opportunities for multiple species to persist in the long
term and, therefore, their coexistence (Godoy et al., 2018). The food
web for the MPA N-BB presents an asymmetric degree distribution. This
pattern was identified at different levels of analysis: food web,
predator, prey, and functional group. Such asymmetry is a well-known
feature in empirical complex food webs in particular (Dunne et al.,
2002a; Montoya \& Solé, 2003; Stouffer et al., 2005), and has received
great attention in complex networks in general (Albert \& Barabási,
2002; Newman, 2003). The degree distribution affects the resilience of
complex food webs against random failures and pressure on a particular
component of the web: food webs showing right-skewed distributions, like
the one described in this study, are more vulnerable to the removal of
the most connected species or hubs, with the potential of producing
secondary extinctions and a catastrophic fragmentation of the network
(Albert et al., 2000; Dunne et al., 2002b; Eklöf \& Ebenman, 2006).

@@

On the other hand, a low number of prey are consumed by many predators
in the food web of the MPA N-BB. This suggests that there are dominant
food sources from which most consumers depend on and from where the
ecosystem energy is being transferred to the upper trophic levels. The
most demanded source we identified in this study (i.e.~detritus)
supports the abundant benthic community of filter-feeders (Schejter et
al., 2016). Furthermore, we found that the second and third-most
consumed prey were diatoms and species of \emph{Euphausia},
respectively, which are essential sources for the diverse zooplankton
community (Spinelli et al., 2020), mid-TL consumers like the Fuegian
sprat \emph{Sprattus fuegensis} (Padovani et al., 2021) and
\emph{Patagonotothen ramsayi} (Fischer et al., 2022), and top predators
(Raya Rey, 2022; Riccialdelli, 2022).

@@

Closeness and betweenness are defined as ``mesoscale'' properties
because they consider direct and indirect interactions, therefore
describing the focal species' ability to influence the rest of the
species of the food web (Lai et al., 2012). Closeness quantifies how
many steps away species \(i\) is from all other species in the food web,
and is proportional to how rapidly the indirect effects of the focal
species can spread to other species in the network (Scotti \& Jordán,
2010). In the food web of the MPA N-BB, low-TL consumers arise as
important in this regard: species of the zooplankton community,
\emph{Calanus} and \emph{Euphausia}, and Brachiopoda. Any perturbation
affecting these species, such as the recently confirmed contaminants
mercury (Fioramonti et al., 2022) and microplastics (Cossi et al., 2021;
Di Mauro et al., 2022) for the area, should be of concern since it might
reach many other species in the food web. Otherwise, betweenness
measures the number of shortest paths between species, providing
information on the importance of species as ``bridges'' for energy
transfer: a species with high betweenness takes part in more food chains
and therefore affects more energy flows (Scotti \& Jordán, 2010). We
have identified the longtail southern cod \emph{Patagonotothen ramsayi}
as the most important species in this sense. Moreover, in light of our
analysis, other species like the Patagonian toothfish \emph{Dissostichus
eleginoides}, the Patagonian cod \emph{Salilota australis}, the
yellowfin notothenioid \emph{Patagonotothen guntheri}, and the
Patagonian longfin squid \emph{Doryteuthis gahi} should be considered as
relevant in the energy transfer in the ecosystem. All these species have
a mid-trophic position in the food web, supporting the `wasp-waist'
control hypothesis for the MPA N-BB (Riccialdelli et al., 2020).

Ecosystems with a pronounced `wasp-waist' structure are suggested to
present a high trophic redundancy, since many species would show similar
trophic habits (Cury et al., 2000). The significant negative
relationship between trophic similarity and trophic level enhances the
hypothesis of functional similarity at low and mid-TL species compared
to higher TL species for the MPA N-BB food web (Riccialdelli et al.,
2020). At the same time, our results highlight the uniqueness in terms
of the trophic role of high-TL predators. Here, not only the expected
pelagic animals such as marine mammals and sea birds arise as relevant,
but also benthic vertebrate (chondrichthyans \emph{Bathyraja macloviana}
and \emph{Squalus acanthias}) and invertebrate species (echinoderms
\emph{Diplopteraster clarki} and \emph{Pteraster sp}) are noteworthy.
The role that such species play in the MPA N-BB ecosystem is unique and
perturbations on them might result in unprecedented changes at the
trophic structure and functioning level.

@@

The food web studied in the present work might be more representative of
the MPA Namuncurá - Banco Burdwood I ecosystem, which comprises the
shallow submarine plateau called Burdwood Bank. This is related to the
sampling effort that was conducted during the research cruises in the
former MPA compared to the MPA N-BB II (i.e.~deep flanks to the south).
As a consequence most of the data we used to build the network come from
studies performed in the MPA N-BB I. Despite this fact, we decided to
build the food web considering both MPAs due to the tight oceanographic
and ecological connection that exists among them (references in APN,
2022).

It's worth mentioning that we did not consider quantitative data
(i.e.~abundance, biomass) to assess the species' role in the food web.
Although there exists such data for some species (Schejter \& Albano,
2021), it would not be possible to include it in the food web framework
described here due to a taxonomical resolution mismatch. In this regard,
we should mention the case of \emph{Zygochlamys patagonica} (Bivalvia)
that is highlighted by our species-level analyses though it has been
found in low abundances in the area (Schejter \& Albano, 2021).

Some species of sessile suspension feeders in high-latitude marine
ecosystems, such as sponges, ascidians and octocorals, avoid predation
by producing secondary metabolites that function as a chemical defense
(Moles et al., 2015; Núñez-Pons et al., 2010; Prieto et al., 2022).
Although this was not yet recorded at the MPA N-BB, there are a few
studies that reported it in other locations in species that inhabit the
MPA N-BB (Rojo de Almeida et al., 2010).

The MPA N-BB I presents complex oceanographic conditions that generate
an internal spatial heterogeneity, mainly along its longitudinal axis
(Matano et al., 2019). Up to date this heterogeneity has been reflected
in phytoplankton and zooplankton communities (Bértola et al., 2018;
García Alonso et al., 2020; Spinelli et al., 2020), and in fish
assemblages (Delpiani et al., 2020). Moreover, seasonal variations also
occur in some physical and biological aspects of the MPA N-BB I (García
Alonso et al., 2018; Matano et al., 2019). Although we were aware of it,
we decided to characterize a food web representing the whole MPA N-BB I
year round since this is the first study of its type in the area.

@@

Incorporate habitat heterogeneity, since the benthic community in the BB
and deep flanks differs {[}; Schejter \& Albano (2021){]}. This
heterogeneity might lead to distinct food web properties, hence
differences in ecosystem functioning and response to anthropogenic and
environmental perturbations (Cordone et al., 2020; Kortsch et al.,
2019). El caso de las ascidias (Tatián)

Incorporate species traits, like body size and mass, which are known to
be important drivers in predator-prey interactions. La Mesa et al.
(2016)

Simulate perturbations on food web and target species: mercury transfer,
microplastic pollution, fisheries. Cossi et al. (2021); Di Mauro et al.
(2022); Fioramonti et al. (2022)

Estimation of interaction strength considering species and interaction
traits (i.e.~body size, body mass, interaction dimensionality) to gain
insights into food web stability and response to anthropogenic and
environmental perturbations. Nilsson \& McCann (2016); Cirtwill et al.
(2018); Kortsch et al. (2021)

@@

\leavevmode\vadjust pre{\hypertarget{ref-Bertola2018}{}}%
Bértola, G., Olguín Salinas, H., \& Alder, V. A. (2018). Distribución
espacial de {Rhizosolenia} crassa, \textquestiondown especie clave del
banco burdwood? \emph{Libro de Resúmenes {X Jornadas Nacionales} de
{Ciencias} Del {Mar}}.
